---
title: "PRACTICAL MACHINE LEARNING -EL HASSANE-"
output: html_document

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## INTRODUCTION 
this is my final project in practical machine learning giving by johnhopkins university in coursera <https://www.coursera.org/specialization/jhudatascience/>
it was about learning the essentials of machine learning algorithmes using caret package <http://caret.r-forge.r-project.org/> 
i used R v 3.2 and Rstudio 1.0.44 as an IDE.

# ABSTRACT
The goal of the  project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).


# INSTALL PACKAGES

```{r cars}
#to use the library  :
library(caret)



#to use the library  :
training <-read.csv(file="C:\\Users\\elhassane\\Desktop\\pml-training.csv")
testing <- read.csv(file="C:\\Users\\elhassane\\Desktop\\pml-testing.csv")
#to have an idea about our data first
summary(training)
set.seed(1000)

## Make  Partition de Data

inTrain<- createDataPartition(y=training$classe,p=0.65,list = FALSE)
traini <- training[inTrain,]
testi <-  training[-inTrain,]
#show dimensions of the partiton
 dim(traini)
 
 #clean the dataser from variables with no change 
nonZero <-nearZeroVar(traini,saveMetrics = TRUE)
nonzerovariables <- names(traini) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")

traini <-traini[!nonzerovariables]
dim(traini)
#delete user_name column and ID column because they have no iinfluence in the prediction
#cleaning data from NA and username
no_names_ID<-c("user_name")
traini <-traini[!names(traini) %in% no_names_ID]
trainingF <- traini 
for(i in 1:length(traini)) { 
        if( sum( is.na( traini[, i] ) ) /nrow(traini) >= .8 ) { #if n?? NAs > 80% of total observations
        for(j in 1:length(trainingF)) {
            if( length( grep(names(traini[i]), names(trainingF)[j]) ) ==1)  { #if the columns are the same:
                trainingF <- trainingF[ , -j] #Remove that column
            }   
        } 
    }
}
dim(trainingF)


# charge the models 

 
library(rpart)
model <-rpart(classe~.,data = trainingF,method="class")
# make prediction 
predictions <-predict(model,testi,type="class")


 #to see accuracy and kappa
 
 #confusionMatrix(predictiions,testi$classe)


```
now apply the prediction to pml-testing


```{r pressure, echo=TRUE}
predictionsfinal <-predict(model,testing,type="class")



```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
